---
title: "05 - Analyses for Contingent Looking Difference"
author: "Janet Bang"
date: "August 19, 2019"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
```

This code includes the analyses for contingent looking.

## Load libraries and functions
```{r}
library(tidyverse)
library(openxlsx)
library(lme4)
library(influence.ME)
library(emmeans)
library(lmerTest)
library(compute.es)
library(lattice)
library(pwr)
library(sjPlot)


theme_set(theme_bw())

source("./helper_functions/R_helperFunctions_WLdata.R")
```


# Contingent looking
## Read in csv files
```{r}
contlook_wl_long <- read_csv("./data/contlook_wl_long.csv") 
demographics <- read.csv("./data/demographics.csv")
```


## Examine the raw number of contingent looks to target and distractor
```{r}
# subset values of number of contingent looks (i.e., summary scores, not difference scores)
# in cue_match, "match" refers to when the cue condition matches the cue in contlook_type
contlook_sum <- contlook_wl_long %>% 
  mutate(motion = factor(motion, c("static", "function"))) %>% 
  mutate(cueCond = factor(cueCond, c("eye", "arw"))) %>% 
  mutate(contlook_type = factor(contlook_type, c("targTeye", "targTarw", "compTeye", "compTarw"))) %>% 
  filter(diff_or_sum == "contlook_sum") %>% 
  filter(cue_match == "match")

# boxplot of means
ggplot(contlook_sum, aes(x = contlook_type, y = value, fill = cueCond)) + 
  geom_boxplot() + 
  facet_wrap(~ motion + group, ncol = 2) + 
  geom_jitter(alpha = .3) + 
  theme(text = element_text(size=20)) +
  labs(y = "Number of Contingent Looks")
ggsave("./figures/contlook_total_looks.pdf", width = 11.5, height = 8, unit = "in")


# descriptives
contlook_sum %>% 
  group_by(participant, group, motion, contlook_type) %>% 
  mutate(mean_part = mean(value), 
         mdn_part = median(value),
         sd_part = sd(value), 
         min_part = min(value), 
         max_part = max(value)) %>% 
  distinct(participant, group, motion, mean_part, mdn_part, min_part, max_part, contlook_type) %>% 
  group_by(group, motion, contlook_type) %>% 
  mutate(mean_group = mean(mean_part), 
         mdn_group = median(mdn_part),
         min_group = min(min_part), 
         max_group = max(max_part)) %>% 
  distinct(group, motion, contlook_type, mean_group, mdn_group, min_group, max_group)
```



## Add language information and select variables for difference scores
```{r}
# select language variable
language <- demographics %>% 
  select(c(participant, language, age_yrs))

# add to contlook df
contlook <- contlook_wl_long %>% 
  left_join(language, by = "participant") %>% 
  filter(diff_or_sum == "contlook_diff")
```


## Exclude rows where cue condition does not match the aoi cue
```{r}
contlook <- contlook %>% 
  filter(cue_match == "match")
```


## Transform variables and re-order levels
```{r}
contlook <- contlook %>% 
  mutate(participant = factor(participant), 
         object = factor(object), 
         trial_num = factor(trial_num), 
         motion = factor(motion), 
         group = factor(group),
         block = factor(block), 
         cueCond = factor(cueCond), 
         video_phase = factor(video_phase), 
         language = factor(language)) %>% 
  mutate(motion = fct_relevel(motion, "static", "function"),
         group = fct_relevel(group, "TYP", "ASD"), 
         cueCond = fct_relevel(cueCond, "eye", "arw"))
```




## Visual analysis: distribution of DV
```{r}
# difference score
ggplot(contlook, aes(x = value, fill = cueCond)) + 
  geom_density(alpha = .3) + 
  facet_wrap(~ motion) 
```



## Mixed models: full group of 41 TD and 22 ASD
```{r}
# comparing models with and without slope
m_contlook = lmerTest::lmer(value ~ cueCond * motion * group + 
                        (1 | participant),
                        data = contlook, REML = F) 

m_contlook_withslope = lmerTest::lmer(value ~ cueCond * motion * group + 
                        (1 + cueCond | participant),
                        data = contlook, REML = F) 

anova(m_contlook, m_contlook_withslope)



# final model for p values calculated with Kenward-Roger's method
# used model without slope bc with slope the corr between slope and intercept is -1
m_contlook_REML = lmerTest::lmer(value ~ cueCond * motion * group + 
                        (1 | participant),
                        data = contlook) 

anova(m_contlook_REML, type = c("III"), ddf = "Kenward-Roger")



# emmeans
object1 <- emmeans(m_contlook_REML, ~ cueCond | motion)
object2 <- emmeans(m_contlook_REML, ~ motion | cueCond)

object1
rbind(pairs(object1), pairs(object2), adjust = "tukey")

# visualization of interaction
emmip(m_contlook_REML, motion ~ cueCond, CIs = T)




# confidence interval of cue comparison: estimate +/- 1.96*SE
# static: gaze - arrow
0.4841611 - (1.96*(0.1682464))
0.4841611 + (1.96*(0.1682464))

# function: gaze - arrow
-0.1237775 - (1.96*(0.1794189))
-0.1237775 + (1.96*(0.1794189))


# Cohen's d for comparison of gaze-arrow
summarySEwithin(contlook, measurevar = "value", 
                withinvars = c("motion", "cueCond"),
                idvar = "participant", na.rm = T, conf.interval=.95)

# static: gaze - arrow
n <- c(122, 119)
mean <- c(1.0000000, 0.5630252)
sd <- c(1.670351, 1.303141)
df <- data.frame(n,mean,sd)
pooled.sd1 <- sqrt(sum(df$sd^2 * (df$n - 1)) / (sum(df$n - 1)) )
mes2(mean[1], mean[2], pooled.sd1, n[1], n[2])


# function: gaze - arrow
n <- c(111, 112)
mean <- c(0.2522523, 0.2946429 )
sd <- c(1.123295, 1.078559)
df <- data.frame(n,mean,sd)
pooled.sd1 <- sqrt(sum(df$sd^2 * (df$n - 1)) / (sum(df$n - 1)) )
mes2(mean[1], mean[2], pooled.sd1, n[1], n[2])



# power for difference between gaze and arrow in static videos
# based on a paired t-test and cohen's d using a pooled sd
pwr.t.test(n = 63, d = .29, sig.level = 0.05, power = NULL, 
    type = c("paired"),
    alternative = c("two.sided"))


# MODEL DIAGNOSTICS
plot_model(m_contlook_REML, type = "diag")



# plot of individual points
ggplot(contlook, aes(x = cueCond, y = value, shape = group, color = group)) + 
  geom_jitter(alpha = .6) + 
  facet_wrap( ~ motion + video_phase) + 
  scale_shape_manual(values = c(19, 17))



# plot for publication
# create values
values <- with(summary(emmeans(m_contlook_REML, ~ cueCond * motion * group)), 
               cbind(cueCond, motion, group, emmean, lower.CL,upper.CL))
values2 <- data.frame(setNames(as.data.frame(values), c("cueCond", "motion", "Group", "mean", "lwr","upr")))
values2 <- transform(values2,
                     motion = factor(motion),
                     cueCond = factor(cueCond), 
                     Group = factor(Group))
values2$cueCond <- revalue(values2$cueCond, c("1" = "gaze", "2" = "arrow"))
values2$motion <- revalue(values2$motion, c("1" = "Static", "2" = "Function"))
values2$Group <- revalue(values2$Group, c("1" = "TD", "2" = "ASD"))


# plot
ggplot(values2, aes(cueCond, y = mean, ymin = lwr, ymax = upr, linetype = Group, shape = Group))+
  geom_line(aes(group = Group), size = 1, position = position_dodge(width = .2)) +
  geom_pointrange(data = values2, position=position_dodge(width = .2), size = 1) + 
  geom_errorbar(width = .3, size = 1, position = position_dodge(width = .2)) +
  labs(y = "Contingent Looking Difference", x = "Cue Condition") + 
  theme_bw(base_size = 12, base_family = "") + 
  theme(text = element_text(size = 30)) + 
  scale_shape_manual(values = c(19, 17)) + 
  geom_hline(yintercept = 0) +
  facet_wrap(~ motion, ncol = 2)
ggsave("./figures/contlook_diff.pdf", width = 12, height = 9, unit = "in")
```

## Visual analysis: language, block to look at cue condition effects
```{r}
# language 
ggplot(contlook, aes(cueCond, value, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ language + video_phase, ncol = 2) 

# block
ggplot(contlook, aes(cueCond, value, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ block + video_phase, ncol = 2) 
```