---
title: "06 - Analyses for target advantage"
author: "Janet Bang"
date: "August 19, 2019"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
```


This code includes the analyses for target advantage scores.


## Load libraries and functions
```{r}
library(tidyverse)
library(openxlsx)
library(lme4)
library(influence.ME)
library(emmeans)
library(lmerTest)
library(compute.es)
library(lattice)
library(pwr)
library(sjPlot)

theme_set(theme_bw())

source("./helper_functions/R_helperFunctions_WLdata.R")
```



# Proportion fixation duration
## Read in csv files
```{r}
fix_dur_wide <- read_csv("./data/fix_dur_wide.csv") 
demographics <- read.csv("./data/demographics.csv")
```


## Add language information and select variables for analysis
```{r}
# select language variable
language <- demographics %>% 
  select(c(participant, language))

# select variables for prop_cue
target_advtg <- fix_dur_wide %>% 
  select(c(participant, object, trial_num, motion, 
           group, seq, block, cueCond, 
           base_targAdvtg, cue1and2sum_targAdvtg,test_targAdvtg)) %>% 
  left_join(language, by = "participant")
```


## Turn to long and create variable names
```{r}
# turn to long
target_advtg_full <- target_advtg %>% 
  gather(variable, value, base_targAdvtg, cue1and2sum_targAdvtg,test_targAdvtg)

# create variables
target_advtg_full <- target_advtg_full %>% 
  mutate(video_phase = ifelse(grepl("base", target_advtg_full$variable), "base", 
                       ifelse(grepl("test", target_advtg_full$variable), "test", "cue1and2sum")), 
         variable = "target_advtg",
         aoi_cue = ifelse(grepl("eye",  target_advtg_full$variable), "eye", "arw")) %>% 
  mutate(video_phase = plyr::revalue(video_phase, c("cue1and2sum" = "teaching"))) %>% 
  select(-variable)
```



## Transform variables and re-order levels
```{r}
target_advtg_full <- target_advtg_full %>% 
  mutate(participant = factor(participant), 
         object = factor(object), 
         trial_num = factor(trial_num), 
         motion = factor(motion), 
         group = factor(group),
         block = factor(block), 
         cueCond = factor(cueCond), 
         video_phase = factor(video_phase), 
         language = factor(language)) %>% 
  mutate(motion = fct_relevel(motion, "static", "function"),
         group = fct_relevel(group, "TYP", "ASD"), 
         cueCond = fct_relevel(cueCond, "eye", "arw"))
```



## Visual analysis: distribution of DV
```{r}
ggplot(target_advtg_full, aes(x = value, fill = cueCond)) + 
  geom_density(alpha = .3) + 
  facet_wrap(~ motion + video_phase) 
```


## Create dataframes for matched groups
```{r}
target_advtg_match <- target_advtg_full %>% 
  filter(participant == 302 | participant == 304 | participant == 306 | participant == 310 |
         participant == 312 | participant == 314 | participant == 315 | participant == 318 |
         participant == 321 | participant == 322 | participant == 323 | participant == 324 |
         participant == 325 | participant == 326 | participant == 329 | participant == 331 |
         participant == 332 | participant == 333 | participant == 341 | participant == 347 |
         participant == 349 | participant == 402 | participant == 403 | participant == 404 |
         participant == 406 | participant == 407 | participant == 408 | participant == 409 |
         participant == 410 | participant == 412 | participant == 413 | participant == 414 |
         participant == 415 | participant == 416 | participant == 418 | participant == 420 |
         participant == 421 | participant == 424 | participant == 425 | participant == 427 |
         participant == 428 | participant == 429 | participant == 430 | participant == 433 |
         participant == 435 | participant == 439 | participant == 443 | participant == 445)

# this results in 24 td and 22 asd (2 asd dropped from eye tracking)
target_advtg_match %>% 
  distinct(participant,group) %>% 
  group_by(group) %>% 
  dplyr::count()
```

## Mixed models: full group of 41 TD and 22 ASD
```{r}
# comparing models with and without slope
m_targAdvtg_full = lmerTest::lmer(value ~ cueCond * video_phase * motion * group + 
                        (1 | participant),
                        data = target_advtg_full, REML = F) 

m_targAdvtg_full_withslope = lmerTest::lmer(value ~ cueCond * video_phase * motion * group + 
                        (1 + cueCond | participant),
                        data = target_advtg_full, REML = F) 

anova(m_targAdvtg_full, m_targAdvtg_full_withslope)



# final model for p values calculated with Kenward-Roger's method
m_targAdvtg_full_REML = lmerTest::lmer(value ~ cueCond * video_phase * motion * group + 
                        (1 + cueCond | participant),
                        data = target_advtg_full) 

anova(m_targAdvtg_full_REML, type = c("III"), ddf = "Kenward-Roger")


# emmeans
object1 <- emmeans(m_targAdvtg_full_REML, ~ video_phase | motion)
object2 <- emmeans(m_targAdvtg_full_REML, ~ motion | video_phase)

object1
rbind(pairs(object1), pairs(object2), adjust = "tukey")

# visualization of interactions
emmip(m_targAdvtg_full_REML, ~ video_phase | motion, CIs = T)

emmip(m_targAdvtg_full_REML, ~ group ~ cueCond | motion, CIs = T)
ggsave("./figures/targAdvtg_3way_n63.pdf")



# confidence interval of cue comparison: estimate +/- 1.96*SE
object3 <- emmeans(m_targAdvtg_full_REML, ~ cueCond)
pairs(object3)

# gaze - arrow
-0.009347344 - (1.96*(0.03730354))
-0.009347344 + (1.96*(0.03730354))


# Cohen's d for cue comparison of gaze-arrow
summarySEwithin(target_advtg_full, measurevar = "value", 
                withinvars = c("cueCond"),
                idvar = "participant", na.rm = T, conf.interval=.95)

n <- c(685, 683)
mean <- c(0.1056441, 0.1297186)
sd <- c(0.7172948, 0.7024318)
df <- data.frame(n,mean,sd)
pooled.sd1 <- sqrt(sum(df$sd^2 * (df$n - 1)) / (sum(df$n - 1)) )
mes2(mean[1], mean[2], pooled.sd1, n[1], n[2])



# MODEL DIAGNOSTICS
plot_model(m_targAdvtg_full_REML, type = "diag")


# plot for publication
# create values
values <- with(summary(emmeans(m_targAdvtg_full_REML, ~ motion * group * video_phase)), 
               cbind(motion, group, video_phase, emmean, lower.CL,upper.CL))
values2 <- data.frame(setNames(as.data.frame(values), c("motion", "Group", "video_phase", "mean", "lwr","upr")))
values2 <- transform(values2,
                     motion = factor(motion),
                     Group = factor(Group), 
                     video_phase = factor(video_phase))
values2$motion <- revalue(values2$motion, c("1" = "Static", "2" = "Function"))
values2$Group <- revalue(values2$Group, c("1" = "TD", "2" = "ASD"))
values2$video_phase <- revalue(values2$video_phase, c("1" = "Baseline", "2" = "Teaching \n(Cue Shift)", "3" = "Test"))


# plot
ggplot(values2, aes(video_phase, y = mean, ymin = lwr, ymax = upr, linetype = Group, shape = Group))+
  geom_line(aes(group = Group), size = 1, position = position_dodge(width = .2)) +
  geom_pointrange(data = values2, position=position_dodge(width = .2), size = 1) + 
  geom_errorbar(width = .3, size = 1, position = position_dodge(width = .2)) +
  labs(y = "Target Advantage", x = "Video Phase") + 
  theme_bw(base_size = 12, base_family = "") + 
  theme(text = element_text(size = 30)) + 
  scale_shape_manual(values = c(19, 17)) + 
  geom_hline(yintercept = 0) + 
  facet_wrap(~ motion, ncol = 2)
ggsave("./figures/target_advantage.pdf", width = 15, height = 9, unit = "in")
```



## Mixed models - matched groups
```{r}
# comparing models with and without slope
m_targAdvtg_match = lmerTest::lmer(value ~ cueCond * video_phase * group * motion + 
                        (1 | participant),
                        data = target_advtg_match, REML = F) 

m_targAdvtg_match_withslope = lmerTest::lmer(value ~ cueCond * video_phase * group * motion + 
                        (1 + cueCond | participant),
                        data = target_advtg_match, REML = F) 

anova(m_targAdvtg_match, m_targAdvtg_match_withslope)



# final model for p values calculated with Kenward-Roger's method
m_targAdvtg_match_REML = lmerTest::lmer(value ~ cueCond * video_phase * group * motion + 
                        (1 + cueCond | participant),
                        data = target_advtg_match) 

anova(m_targAdvtg_match_REML, type = c("III"), ddf = "Kenward-Roger")


emmip(m_targAdvtg_match_REML, ~ group ~ cueCond | motion, CIs = T)
# ggsave("targAdvtg_3way_n46.pdf")


# MODEL DIAGNOSTICS
plot_model(m_targAdvtg_match_REML, type = "diag")

```


## Visual analysis: language, block to look at potential cue condition effects
```{r}
# language 
ggplot(target_advtg_full, aes(cueCond, value, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ language + video_phase, ncol = 3) 

# block
ggplot(target_advtg_full, aes(cueCond, value, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ block + video_phase, ncol = 3) 
```