---
title: "09 - Analyses for the number of semantic features"
author: "Janet Bang"
date: "August 19, 2019"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE, cache=FALSE)
```


This code includes the analyses for description data.

NOTE: One participant (312) only has 7 total descriptions instead of a possible 8 descriptions (across visits 1 and 2) because 1 was administered incorrectly. 


## Load libraries and functions
```{r}
library(tidyverse)
library(openxlsx)
library(lme4)
library(influence.ME)
library(emmeans)
library(lmerTest)
library(compute.es)
library(lattice)
library(GGally)
library(sjPlot)
library(effects)
library(Hmisc)
library(pwr)

theme_set(theme_bw())

source("./helper_functions/R_helperFunctions_WLdata.R")
```


## Read in csv files
```{r}
descrip_sum <- read.csv("./data/descrip_sum.csv")
exp_info_wl <- read.csv("./data/exp_info_wl.csv")
demographics <- read.csv("./data/demographics.csv")
```


## Combine info in files
```{r}
# Select variables in demographics and combine with description data
demo_descrip_sum <- demographics %>% 
  select(c(participant, age_yrs, language, sex, mother_ed)) %>% 
  right_join(descrip_sum, by = "participant")

# Select variables in exp_info_wl and combine with new description data
descrip <- exp_info_wl %>%
  filter(motion == "static") %>%  
  select(-c(practice, motion, eyetk_inc)) %>% 
  right_join(demo_descrip_sum, by = c("participant", "object"))
```

## Transform variables and reorder levels
```{r}
descrip <- descrip %>% 
  mutate(participant = factor(participant), 
         seq = factor(seq), 
         object = factor(object),
         trial_num = factor(trial_num), 
         block = factor(block), 
         cueCond = factor(cueCond), 
         language = factor(language),
         visit = factor(visit)) %>% 
  mutate(group = fct_relevel(group, "TYP", "ASD"), 
         cueCond = fct_relevel(cueCond, "eye", "arw"))


# check number of participants: ASD n = 22; TD = 43
# 1 child with ASD could not complete the task at all and 2 children with ASD were not able to provide descriptions of both familiar objects at both visits
descrip %>% 
  distinct(participant,group) %>% 
  group_by(group) %>% 
  dplyr::count()
```

# Number of correct target objects
## Create column for binary coding of whether child described target object
```{r}
descrip <- descrip %>% 
  mutate(target_bin = ifelse(correct_object == "target_object", 1, 0))
```



## Frequency of children per group with valid and invalid descriptions
```{r}
sum_descrip <- descrip %>% 
  group_by(participant, group, visit) %>% 
  tally(target_bin) %>% 
  mutate(num_correct = factor(n))

sum_descrip2 <- sum_descrip %>% 
  group_by(group, visit, num_correct) %>% 
  dplyr::count() %>% 
  mutate(total_children = ifelse(group == "TYP", 43, 22)) %>% 
  mutate(prop_children = n / total_children)

ggplot(sum_descrip2, aes(x = num_correct, y = prop_children, fill = group)) + 
  geom_col(position = "dodge") + 
  facet_wrap(~ visit)
```



## Mixed models - everyone
```{r}
# comparing models with and without slope
m_descrip_correct_target = glmer(target_bin ~ visit * group + 
                                   (1 | participant), data = descrip, family = binomial,
                                   control = glmerControl(optimizer="bobyqa")) 
m_descrip_correct_target_withslope = glmer(target_bin ~ visit * group + 
                                   (1 + cueCond | participant), data = descrip, family = binomial,
                                   control = glmerControl(optimizer="bobyqa")) 

anova(m_descrip_correct_target, m_descrip_correct_target_withslope)

# final model for analysis
summary(m_descrip_correct_target)
car::Anova(m_descrip_correct_target, type = "III", ddf = "Kenward-Roger")


# MODEL DIAGNOSTICS
plot_model(m_descrip_correct_target, type = "diag")
```




# Number of semantic features
## Select only target object descriptions and create summary score of semantic features
```{r}
# Select only target object descriptions
descrip_mm <- descrip %>% 
  filter(correct_object == "target_object") %>% 
  mutate(participant = factor(participant))

descrip_mm <- descrip_mm %>% 
  mutate(sem_features = sum_intd_fxn + sum_phys_attb + sum_imagined_phys_attb)

descrip_mm %>% 
  distinct(participant, group) %>% 
  ungroup() %>% 
  group_by(group) %>% 
  tally()
```


## Visual analysis: distribution of DV
```{r}
ggplot(descrip_mm, aes(x = sem_features, fill = cueCond)) + 
  geom_density(alpha = .3) + 
  facet_wrap(~ visit) 


# distribution of # of observations from each participant
obs_per_participant <- descrip_mm %>% 
  distinct(participant, group, object, visit, cueCond) %>% 
  group_by(participant, group, cueCond) %>% 
  dplyr::count()

ggplot(obs_per_participant, aes(n, fill = cueCond)) + 
  geom_histogram(position = position_dodge())
```



## Mixed models: full group of 41 TD and 19 ASD (the model excludes children who never provided any valid descriptions)
```{r}
# comparing models with and without slope
m_descrip = lmerTest::lmer(sem_features ~ cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm, REML = F) 

m_descrip_withslope = lmerTest::lmer(sem_features ~ cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm, REML = F) 

anova(m_descrip, m_descrip_withslope)



# final model for p values calculated with Kenward-Roger's method
m_descrip_REML = lmerTest::lmer(sem_features ~ cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm) 

anova(m_descrip_REML, type = c("III"), ddf = "Kenward-Roger")



# confidence interval of cue comparison: estimate +/- 1.96*SE
object1 <- emmeans(m_descrip_REML, ~ cueCond)
object1
pairs(object1)

# gaze - arrow
0.6961047 - (1.96*(0.3741881))
0.6961047 + (1.96*(0.3741881))


# Cohen's d for cue comparison of gaze-arrow
summarySEwithin(descrip_mm, measurevar = "sem_features", 
                withinvars = c("cueCond"),
                idvar = "participant", na.rm = T, conf.interval=.95)

n <- c(121, 123)
mean <- c(6.148760 , 5.707317)
sd <- c(3.159950, 2.781065)
df <- data.frame(n,mean,sd)
pooled.sd1 <- sqrt(sum(df$sd^2 * (df$n - 1)) / (sum(df$n - 1)) )
mes2(mean[1], mean[2], pooled.sd1, n[1], n[2])




# MODEL DIAGNOSTICS
plot_model(m_descrip_REML, type = "diag")
```



## Visual analysis: language, block to look at cue condition effects
```{r}
# language 
ggplot(descrip_mm, aes(cueCond, sem_features, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ language, ncol = 2) 

# block
ggplot(descrip_mm, aes(cueCond, sem_features, fill = cueCond)) + 
  geom_boxplot() + 
  geom_jitter(alpha = .3) + 
  stat_summary(fun.data = give.n, geom = "text", size = 6, position=position_dodge(width=.9)) +
  stat_summary(fun.y = mean, shape = 16, col='red', geom = 'point', size = 3, position=position_dodge(width=.9)) + 
  facet_wrap(~ block, ncol = 2) 
```



# Relationships with covariates - Exploratory
## Read in csv files
```{r}
celf <- read.csv("./data/celf.csv")
leiter <- read.csv("./data/leiter.csv")
vineland <- read.csv("./data/vineland.csv")
```



## Select variables and combine into one df
```{r}
# celf
celf2 <- celf %>%
  dplyr::select(c(participant, celf4_RS_scaled, celf4_WC_total_scaled, celf4_WA_total))

# leiter
leiter2 <- leiter %>% 
  dplyr::select(c(participant, leiter_composite)) %>% 
  right_join(celf2, by = "participant")

# vineland
vineland2 <- vineland %>% 
  dplyr::select(c(participant, social_standScore)) %>% 
  right_join(leiter2, by = "participant") %>% 
  mutate(participant = factor(participant))
```


## Combine above covariate info with description data
```{r}
descrip_mm_cov <- descrip_mm %>% 
  left_join(vineland2, by = "participant")

# check number of participants: 19 ASD, 41 TD
descrip_mm_cov %>% 
  distinct(participant,group) %>% 
  group_by(group) %>% 
  dplyr::count()

```



## Visual analysis of Vineland x semantic features
```{r}
ggplot(descrip_mm_cov, aes(social_standScore, sem_features, color = cueCond)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ group)


# removing outlier with high Vineland score (TD participant)
descrip_mm_cov_rm1 <- descrip_mm_cov %>% 
  filter(social_standScore != 160)

ggplot(descrip_mm_cov_rm1, aes(social_standScore, sem_features, color = cueCond)) +
  geom_point() + 
  geom_smooth(method = "lm") + 
  facet_wrap(~ group)
```



## Center covariates
```{r}
centered_scores <- descrip_mm_cov %>% 
  distinct(participant, social_standScore, celf4_RS_scaled, celf4_WC_total_scaled) %>% 
  mutate(vineland_cntr = scale(social_standScore, scale = F)) %>% 
  mutate(vineland_cntr = as.numeric(vineland_cntr)) %>% 
  mutate(celf4_RS_scaled_cntr = scale(celf4_RS_scaled, scale = F)) %>% 
  mutate(celf4_RS_scaled_cntr = as.numeric(celf4_RS_scaled_cntr)) %>% 
  mutate(celf4_WC_total_scaled_cntr = scale(celf4_WC_total_scaled, scale = F)) %>% 
  mutate(celf4_WC_total_scaled_cntr = as.numeric(celf4_WC_total_scaled_cntr))

descrip_mm_cov <- centered_scores %>% 
  right_join(descrip_mm_cov, by = c("participant", "social_standScore", "celf4_RS_scaled", "celf4_WC_total_scaled"))
```


## Mixed models: full group of 41 TD and 19 ASD with Vineland
```{r}
# comparing models with and without slope
m_vineland = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov, REML = F) 

m_vineland_withslope = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm_cov, REML = F)

anova(m_vineland, m_vineland_withslope)


# final model for p values calculated with Kenward-Roger's method
m_vineland_REML = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group +
                        (1 | participant),
                        data = descrip_mm_cov) 

anova(m_vineland_REML, type = c("III"), ddf = "Kenward-Roger")



# confidence interval of cue comparison: estimate +/- 1.96*SE
object1 <- emmeans(m_vineland_REML, ~ cueCond)
object1
pairs(object1)

# gaze - arrow
1.207094 - (1.96*(0.6380356))
1.207094 + (1.96*(0.6380356))


# power for difference between gaze and arrow in static videos
# based on a paired t-test and cohen's d using a pooled sd
pwr.t.test(n = 60, d = .15, sig.level = 0.05, power = NULL, 
    type = c("paired"),
    alternative = c("two.sided"))





# post hoc testing of trends
trends_cueCond <- emtrends(m_vineland_REML, pairwise ~ cueCond, var = "vineland_cntr" )
trends_group <- emtrends(m_vineland_REML, pairwise ~ group, var = "vineland_cntr" )

trends_cueCond
summary(trends_cueCond, infer=c(TRUE,TRUE), null=0)
summary(trends_group, infer=c(TRUE,TRUE), null=0)

# visualization of interactions
emmip(m_vineland_REML, cueCond ~ vineland_cntr, cov.reduce = range)
emmip(m_vineland_REML, group ~ vineland_cntr, cov.reduce = range)


# plots for publication
plot_model(m_vineland_REML, type = "eff", terms = c("vineland_cntr", "cueCond"), 
           color = c("blue", "red"), 
           axis.lim = c(0, 16)) 
ggsave("./figures/sem_features_vineland_slopes.pdf", width = 11.5, height = 8, unit = "in")


ggplot(descrip_mm_cov, aes(x = vineland_cntr, y = sem_features, shape = group, color = cueCond)) + 
  geom_jitter(size = 3) + 
  scale_shape_manual(values = c(1,17)) + 
  ylim(0, 16) + 
  scale_color_manual(values = c("red", "blue"))
ggsave("./figures/sem_features_vineland_points.pdf", width = 11.5, height = 8, unit = "in")



# MODEL DIAGNOSTICS
plot_model(m_vineland_REML, type = "diag")
```




## Mixed models: 40 TD and 19 ASD - removing TD outlier with high Vineland score
```{r}
descrip_mm_cov_rm1 <- descrip_mm_cov %>% 
  filter(social_standScore != 160)

# comparing models with and without slope
m_vineland_rm1 = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov_rm1, REML = F) 

m_vineland_withslope_rm1 = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm_cov_rm1, REML = F)

anova(m_vineland_rm1, m_vineland_withslope_rm1)


# final model for p values calculated with Kenward-Roger's method
m_vineland_REML_rm1 = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov_rm1) 

anova(m_vineland_REML_rm1, type = c("III"), ddf = "Kenward-Roger")


# post hoc testing of trends
trends_cueCond_n59 <- emtrends(m_vineland_REML_rm1, pairwise ~ cueCond, var = "vineland_cntr" )

trends_cueCond_n59
summary(trends_cueCond_n59, infer=c(TRUE,TRUE), null=0)

# plots
emmip(m_vineland_REML_rm1, cueCond ~ vineland_cntr, cov.reduce = range)



# MODEL DIAGNOSTICS
plot_model(m_vineland_REML_rm1, type = "diag")
```



## Create dataframes for full TD only and matched groups
```{r}
descrip_mm_cov_match <- descrip_mm_cov %>% 
  filter(participant == 302 | participant == 304 | participant == 306 | participant == 310 |
         participant == 312 | participant == 314 | participant == 315 | participant == 318 |
         participant == 321 | participant == 322 | participant == 323 | participant == 324 |
         participant == 325 | participant == 326 | participant == 329 | participant == 331 |
         participant == 332 | participant == 333 | participant == 341 | participant == 347 |
         participant == 349 | participant == 402 | participant == 403 | participant == 404 |
         participant == 406 | participant == 407 | participant == 408 | participant == 409 |
         participant == 410 | participant == 412 | participant == 413 | participant == 414 |
         participant == 415 | participant == 416 | participant == 418 | participant == 420 |
         participant == 421 | participant == 424 | participant == 425 | participant == 427 |
         participant == 428 | participant == 429 | participant == 430 | participant == 433 |
         participant == 435 | participant == 439 | participant == 443 | participant == 445)

# this results in 23 asd and 18 td (these are children who have valid descriptions of the target)
descrip_mm_cov_match %>% 
  distinct(participant,group) %>% 
  group_by(group) %>% 
  dplyr::count()
```



## Mixed models: matched group 23 TD and 18 ASD
```{r}
# comparing models with and without slope
m_vineland_matched = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov_match , REML = F) 

m_vineland_withslope_matched = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm_cov_match , REML = F)

anova(m_vineland_matched, m_vineland_withslope_matched)



# final model for p values calculated with Kenward-Roger's method
m_vineland_REML_matched = lmerTest::lmer(sem_features ~ vineland_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov_match) 

anova(m_vineland_REML_matched, type = c("III"), ddf = "Kenward-Roger")



# MODEL DIAGNOSTICS
plot_model(m_vineland_REML_matched, type = "diag")
```




## Correlation matrix for covariates
```{r}
corr_mtx_cov <- descrip_mm_cov %>% 
  select(c(age_yrs, leiter_composite, 
           celf4_RS_scaled, celf4_WC_total_scaled, celf4_WA_total, social_standScore)) %>% 
  distinct()

rcorr(as.matrix(corr_mtx_cov), type = "pearson")

# plot different comparisons
ggplot(corr_mtx_cov, aes(celf4_WA_total, social_standScore)) + 
  geom_point() + 
  geom_smooth(method = "lm")

```



## Mixed models: full group of 41 TD and 19 ASD with celf4_WC_total
```{r}
# comparing models with and without slope
m_celf4_WC = lmerTest::lmer(sem_features ~ celf4_WC_total_scaled_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov, REML = F) 

m_celf4_WC_withslope = lmerTest::lmer(sem_features ~ celf4_WC_total_scaled_cntr * cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm_cov, REML = F)

anova(m_celf4_WC, m_celf4_WC_withslope)


# final model for p values calculated with Kenward-Roger's method
m_celf4_WC_REML = lmerTest::lmer(sem_features ~ celf4_WC_total_scaled_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov) 

anova(m_celf4_WC_REML, type = c("III"), ddf = "Kenward-Roger")


# MODEL DIAGNOSTICS
plot_model(m_celf4_WC_REML, type = "diag")
```




## Mixed models: full group of 41 TD and 19 ASD with celf4_RS_scaled
```{r}
# comparing models with and without slope
m_celf4_RS = lmerTest::lmer(sem_features ~ celf4_RS_scaled_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov, REML = F) 

m_celf4_RS_withslope = lmerTest::lmer(sem_features ~ celf4_RS_scaled_cntr * cueCond * visit * group + 
                        (1 + cueCond | participant),
                        data = descrip_mm_cov, REML = F)

anova(m_celf4_RS, m_celf4_RS_withslope)


# final model for p values calculated with Kenward-Roger's method
m_celf4_RS_REML = lmerTest::lmer(sem_features ~ celf4_RS_scaled_cntr * cueCond * visit * group + 
                        (1 | participant),
                        data = descrip_mm_cov) 

anova(m_celf4_RS_REML, type = c("III"), ddf = "Kenward-Roger")


# MODEL DIAGNOSTICS
plot_model(m_celf4_RS_REML, type = "diag")
```

